---
title: 02 - Sad Plot Made Better
author: Saundra Schlesinger
date: '2018-06-10'
slug: 02-sad-plot-made-better
categories:
  - R
tags: []
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include = FALSE}
library(here)
library(tidyverse)
library(extrafont)
```

CHALLENGE #3
```{r}
#Read in new dataset, post-1981 males only with post-IFOCE TRUE/FALSE column
hdm_affil <- read_csv(here::here("Week1Lab", "hot_dog_contest_with_affiliation.csv"), 
    col_types = cols(
      affiliated = col_factor(levels = NULL), 
      gender = col_factor(levels = NULL)
      )) %>% 
  mutate(post_ifoce = year >= 1997) %>% 
  filter(year >= 1981 & gender == "male") 
glimpse(hdm_affil)
```

CHALLENGE #4
```{r}
#Use dplyr::distinct to figure out how many unique values there are of affiliated.
#Use dplyr::count to count the number of rows for each unique value of affiliated; use ?count to figure out how to sort the counts in descending order.
hdm_affil %>%
  distinct(affiliated)

hdm_affil %>% 
  count(affiliated, sort=TRUE)
```

CHALLENGE #5
```{r}
#Update the colors using hex colors: c('#E9602B','#2277A0','#CCB683').
#Change the legend title to "IFOCE-affiliation".
#Save this plot object as "affil_plot".
affil_plot<- ggplot(hdm_affil, aes(x = year, y = num_eaten, fill=affiliated)) + #fill by IFOCE status
  geom_col(colour = "white") + #make all columns outlined in white
  labs(title = "Nathan's Hot Dog Eating Contest Results, 1981-2017", x = "Year", y = "Hot Dogs and Buns Consumed")+
  scale_fill_manual(values=c("#E9602B", "#2277A0", "#CCB683"), name="IFOCE-affiliation") #change colors and legend title
```

FINAL FREE-FOR-ALL PLOT

Get female data
```{r}
hdm_females <- read_csv(here::here("data", "hot_dog_contest_with_affiliation.csv"), 
    col_types = cols(
      affiliated = col_factor(levels = NULL), 
      gender = col_factor(levels = NULL)
      )) %>% 
  mutate(post_ifoce = year >= 1997) %>% 
  filter(year >= 1981 & gender == "female") 
```


```{r}
affil_plot +
  geom_col(data = hdm_females, width =0.75, fill = "#92ebb1") + #add in female data
  annotate('text', x = 2018.3, y = 39, label="*", size = 6) +
  labs(caption = "* From 2011 on, separate Men's and Women's prizes have been awarded. \n All female champions to date have been MLE/IFOCE-affiliated.") + #add caption
  annotate('text', x=1992, y=55.5, label="Kobayashi revolutionizes the game with novel technique, \nleading to a doubling of the previous year's best", size = 3, color="#d95f02") +
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 80, 10)) +
  scale_x_continuous(expand = c(0, 0))+
  coord_cartesian(xlim=c(1980,2019), ylim=c(0,80)) +
  theme(plot.title = element_text(hjust = 0.5), #center title
    text=element_text(family="Verdana"),
    axis.text = element_text(size = 10), #set axis font size
    panel.background = element_blank(), #remove grey background
    panel.grid.major.y = element_line(color = "gray95"), #adds back in major gridlines
    panel.border = element_rect(color = "gray80", fill = NA), #creates a border around plot
    axis.line.x = element_line(color = "gray80", size = 0.5), #change size and color of tick marks on x-axis
    axis.ticks = element_line(color = "gray80", size = 0.5), #change size and color of tick marks on y-axis
    legend.background = element_rect(colour = "gray80", fill = "white", size = 0.25, linetype='solid')) +
  scale_fill_manual(values=c("#1b9e77", "#d95f02", "#7570b3"), name="IFOCE-affiliation", guide = guide_legend(reverse = TRUE)) #change bar colors and legend order to match appearance order in graph
```

1) Type of graph:
This is a bar graph.

2) Description of data:
These data were from a list of hot dog eating champions, including related data (like year and number eaten). Also included were gender and IFOCE-affiliation status.

3) Description of audience:
This would be a person interested in the history of hot-dog eating competitions and those interested in Takeru Kobayashi's history in the game.

4) Representation Description:
I am trying to show how dramatically the winning number of hotdogs has increased over time, particularly with the introduction of Takeru Kobayashi to the game.

5) How to read it & What to look for: 
Pay attention to the dramatic difference in 2000. Also pay attention to the number of hot dogs female winners consume versus male competitors.

6) Presentation tips: 
The colors are categorical because although affiliation does have a some ordinality, I did not feel that ordinality needed to be highlighted. The scale is linear so it is easy to appreciate the dramatic differences in winner consumption.

7) Variations and alternatives: 
An alternative would be a line graph. I feel a bar graph better highlights the amounts, whereas a line graph better highlights the changes over time.

8) Methods:
 For this graph, I pulled all the male and female data and plotted based on how much the winner ate each year. I also colored by IFOCE affliation. Finally I added some annotation to add additional context to the graph.

