---
title: 01 - Sad Plot Made Better
author: Saundra Schlesinger
date: '2018-06-10'
slug: 01-sad-plot-made-better
categories:
  - R
tags: []
description: Desc
hacker_news_id: ''
lobsters_id: ''
meta_img: /images/image.jpg
---

Call In Libraries and Data
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(babynames)
library(stringr)
```  
  
Add column of first letters
```{r}
baby_letters <- babynames %>%
  mutate(first_letter=str_sub(name, 1,1))
```

Add up proportion for each letter by year + sex
```{r}
letter_by_year <- baby_letters %>%
  count(year, sex, first_letter, wt = prop) %>% #wt=prop is taking all [year] [letter] names proportions and adding the proportions
  rename(total_prop=nn)
```
  
Label most popular for each year + sex
```{r}
letter_by_year<- letter_by_year %>%
  group_by(sex, year) %>%
  mutate(top_yearly=case_when(
    total_prop == max(total_prop) ~ first_letter)) %>% #think of left of tilde as logical. Ex: if total_prop is equal to max total prop then variable is first letter
  ungroup() %>% #want to return only current year, so need to ungroup
  mutate(top_latest=case_when(
    year == max(year) ~ top_yearly #top yearly is equal to max (current) year
  ))
```
  

Make a data frame with just the latest top letters
```{r}
top_current <- letter_by_year %>%
  filter(!is.na(top_latest))%>% #pull out the not-N/A rows
  rename(color_by = top_latest) %>% 
  select(sex, first_letter, color_by)
```
  

Now merge back in
```{r}
letter_by_year <- letter_by_year %>%
  left_join(top_current) %>%
  mutate(color_by = replace_na(color_by, "else")) #replace N/A with string "else"
```
  
Set up colors for males and females
```{r}
my_colors <- c("#FF1493", "gray80", "#3399ff")
```

Change proportion to percent
```{r}
letter_by_year <- letter_by_year %>%
  mutate(total_prop = total_prop*100)
```
  
Plot it!
```{r}
ggplot(letter_by_year, aes(x = year, y = total_prop, color = color_by, group = first_letter)) +
  geom_line()+
  scale_color_manual("First Letter", values = my_colors) +
  facet_wrap(~sex) +
  ggtitle("First Letter Baby Name Trends", subtitle = "Popularity of first letter of baby names fluctuates over time.") +
  labs(x="Year", y="Proportion of Baby Names Beginning with Letter (%)", family = "Stencil") +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.background = element_rect(fill = "white"),
    axis.title = element_text(face="bold"), 
    title = element_text(face="bold"), 
    axis.text = element_text(size = 10), 
    panel.grid.major.y = element_blank(), 
    panel.grid.major.x = element_blank(), 
    panel.grid.minor.y = element_blank(), 
    panel.grid.minor.x = element_blank(), 
    axis.line.x = element_line(color = "gray30", size = .5), 
    axis.line.y = element_line(color = "gray30", size = .5), 
    axis.ticks = element_line(color = "gray30", size = 1)) 
```
    


Type of graph: This is a line graph.

Description of data: These data are from the babynames set, which tracks incidence of babies' names over time.

Description of audience: These data could be for a number of audiences. It could be for people looking to name their own baby - do they want to be with a trend or not? This could be for indivuduals interested in their own name's letter's popularity. Some letters have shown steady popularity, but others had a period of popularity before falling back off.

Representation Description: I am trying to highlight the current popular first letter, and how its popularity has changed over time. Today's most popular first letter highlighted and then is graphed against all other letters over time, faceted by gender. You can see the percent of names represented on the y-axis, and the year on the x-axis.

How to read it & What to look for: Look how the trend of popularity has changed. For instance, in this case J is currently popular and has been popular through the last century for boys' names. A for girls, though, was only moderately popular before suffering a bout of unpopularity after WWII, but then quickly rebounded in the 60's to the number one spot. Very interesting! Another thing to look for is how the percent representation of the popular letter compares to other letters. For both, the popular letter makes up a little over 15% of names, and the next most popular letter about 10%. That has changed with time, for instance in the boys' case, J has remained most popular by several percentage points over time except for a period around 1950 when another letter came quite close.

Presentation tips: Color was used to emphasize the current popular letter and de-emphasize all other letters. The data are faceted to provide a chart for both boys' and girls' names. The colors chosen are the stereotypical 'boys/blue' and 'girls/pink' drawing more attention to the differences between the two.

Variations and alternatives: A similar chart would be a stacked area graph, but I argue it would be detrimental to the visualization by adding unneeded noise. I think this is the best method and there is no good alternative, really.

Methods: For this graph I pulled out the first letter from all the names in the list over time. I then collected them to determine how many of each letter were represented in each year. The letter with the highest number for the latest year was pulled out as the 'popular' letter. That letter was then graphed against all other letters over time.